\documentclass[12pt]{article}
% \addtolength{\textwidth}{6mm}
% \addtolength{\textheight}{20mm}
%\newcommand{\mmm}{W}
\usepackage{amsfonts,amssymb,amsmath}


\begin{document}

\begin{center}
Notes on polymer adsorption --- revised Aug 15, 2019
\end{center}

\section{Mean-field lattice model}

We shall use a lattice model to describe our experimental situation.  
Lattices will appear in two ways:  (1) covering the surface of a sphere by a two-dimensional lattice, and
(2) filling space with a regular three-dimensional lattice.  
The geometry of the lattices will not be important.  We shall use the simple cubic lattice for our
three-dimensional lattice.  
We shall assume that the two-dimensional lattice has coordination number $q$ 
(that is, each site on the surface has exactly $q$ neighbouring sites).  We assume that 
the distance between lattice sites corresponds to the distance between adjacent monomers within a 
polymer.  We assume that in equilibrium, \textit{every lattice site in the surface is covered by an adsorbed 
monomer}.

Our experiment has a large number $n_{sphere}$ of spheres in a solution of volume $V_{sol}$.  To simplify 
our model and keep the concentrations the same, we shall consider only a single sphere in a solution
of volume $V_1:=V_{sol}/n_{sphere}$.  We shall refer to this as our model system.
We shall view this small three-dimensional region of the solution in our model system as being filled by a 
cubic lattice.
Let $V_{latt}$ be the number of lattice sites in our solution region of volume $V_1$. 

We shall consider polymers of two different lengths, which we shall call ``short'' and ``long.''  Their lengths
will be described by two positive integer parameters $W$ and $n$.

\begin{verse}
A \textbf{short polymer} will consist of $W$ monomers.  
Let $n_S$ be the number of short polymers in the model system. 
\\
A \textbf{long polymer} will consist of $nW$ monomers.  
Let $n_L$ be the number of long polymers in the model system. 
\end{verse}
We shall often refer to our short and long polymers as $W$-mers and $nW$-mers respectively.
We shall want the total mass of the two kinds of monomers in our system to be the same, so we 
shall assume that $n_S \,=\, n\cdot n_L$.

Let $A_{latt}$ be the number of lattice sites on the surface of one sphere, and let $a$
be the number of long polymers that would exactly cover all the sites of the surface:
\[    a  \;=\;  \frac{A_{latt}}{nW}   \,  .  \]
Then $na$ is the number of short polymers that can completely cover the surface.
For simplicity, we assume that $a$ is an integer.

We shall see that the underlying entropic reason that
mostly long chains get adsorbed is that $V_{latt}\gg A_{latt}$, i.e.\ there are many more possible locations in 
the three-dimensional solution than on the two-dimensional surface.


We shall model our polymer configurations in the solution in a slightly different way from those on the surface.
The difference in approach is due to the fact that the polymers are dense on the surface but 
dilute in the solution.  We shall model the polymer configurations in the solution as a collection of 
self-avoiding walks in the cubic lattice, with no interaction between walks (because of the dilute solution).
For the polymers adsorbed densely on the surface lattice, we shall use mean-field calculations 
of the Flory-Huggins type 
% (similar to the lattice approach in Park et al., Macromolecules 2001;
(see also Section XII.1 of \cite{Flo}).

Recall that our ensemble consists of $n_S$ (short) $W$-mers and $n_L$ (long) $nW$-mers.
Of these, some are adsorbed onto the surface of the sphere, and the others float freely in the 
solution of volume $V_1$.  We write $j$ for the number of (long) $nW$-mers that are adsorbed 
onto the surface in a particular configuration of the system.  This implies that
\begin{verse}
 ($i$) the number of $nW$-mers in solution is $n_L-j$;
 \\
 ($ii$) the number of $W$-mers adsorbed on the surface is $n(a-j)$ (because there are $anW$ sites on the
 surface, all occupied, of which $jnW$ are occupied by $nW$-mers, leaving $anW-jnW$ to be 
 occupied by $W$-mers); and
 \\
 ($iii$) the number of $W$-mers in solution is $n_S-n(a-j)$.
 \end{verse}  
We shall write $\mathcal{E}_j$ for the set of all configurations that have exactly $j$ $nW$-mers adsorbed 
onto the surface.  Observe that the possible values of $j$ are the integers from 0 to $a$.
The full space of configurations in our model is $\cup_{j=0}^a {\mathcal{E}_j}$, 
which we shall call $\mathcal{E}$.

Since the only energy in this model comes from the adsorption contacts between monomer and surface,
and since the number of such contacts is the same (namely $A_{latt}$) in every configuration
in $\mathcal{E}$, we see that the energy of every configuration of $\mathcal{E}$ is exactly the same.
Therefore, according to the Boltzmann distribution, \textit{every configuration in $\mathcal{E}$ is equally
likely}.  Thus the calculation of probabilities of events is equivalent to counting the configurations.
In particular, writing $|\mathcal{A}|$ for the number of configurations in a set $\mathcal{A}$, we have
\[
    \hbox{Probability that $j$ long polymers are adsorbed}  \;=\; 
    \frac{|\mathcal{E}_j|}{|\mathcal{E}|}\,,   \hspace{5mm}(j=0,1,\ldots,a).
\]
Thus, to find the most likely number of $nW$-mers (and $W$-mers) to be adsorbed onto the sphere,
we need to find which of the sets $\mathcal{E}_j$ is largest.
We now proceed to estimate the size of each $\mathcal{E}_j$.

\smallskip

The configuration of a single  polymer in a dilute solution is often modeled by
a self-avoiding walk (SAW) in a lattice, i.e.\  a path in the lattice that does not visit any site more than once.
A fundamental property of SAWs is the following.  
Let $c_{\ell}$ be
the number of SAWs that start from a specified lattice site (``the origin'') and visit a total of $\ell$ sites
(including the starting site).   Then the 
asymptotic behaviour of $c_{\ell}$ on the three-dimensional cubic lattice is 
\begin{equation}
    \label{eq.sawscale}
       c_{\ell}  \;\sim  \;  A_3 \, {\ell}^{\gamma_3-1}  \mu_3^{\ell}    \hspace{5mm}\hbox{as $\ell\rightarrow\infty$}.
\end{equation}
% (This has not been proven rigorously, but even mathematicians do not doubt its truth.) 
Here $\mu_3$ and $A_3$ depend on the choice of lattice, but $\gamma_3$ is a universal critical exponent
that is the same for all three-dimensional lattices.  
Their values are known to be approximately \cite{CL,MS}
\begin{equation}
   \label{eq.gammas}   \mu_3 \;=\;  4.684, \hspace{5mm}
        \gamma_3 \;=\;  1.162,    \hspace{5mm}\hbox{and}\hspace{5mm}
    A_3  \;=\;    0.2573  \,.  %1.205/4.683 \;=\;  \,.
\end{equation}
(Estimation in \cite{CL} is for the 
number of SAWs with $\ell$ \textit{steps} (i.e.\ $\ell+1$ \textit{sites}), which we write  $c_{\ell+1}=(A_3\mu_3)\ell^{\gamma_3-1}\mu_3^{\ell}$; reference \cite{CL} obtains $A_3\mu_3=1.205$.)

Since the number of sites in the lattice corresponding to the region of solution is $V_{latt}$, 
we have $V_{latt}c_{\ell}$ ways to place a polymer of size ${\ell}$.  As we are neglecting the
possibility of overlapping polymers in the dilute solution, it follows that the number of ways to 
place $N$ identical $\ell$-mers is 
\begin{equation}
  \label{eq.Npoly}
   \frac{(V_{latt}c_{\ell})^N}{N!}  \,.   
\end{equation}

\smallskip

For a particular choice of $j$, we want to know the number of ways to place $j$ $nW$-mers 
and $a(n-j)$ $W$-mers on the surface, without overlapping, to fill all $anW$ lattice sites on the surface.
This is a hard combinatorial problem, so we shall use the following mean-field (Flory-Huggins) approach.
We shall place one polymer at a time, starting with the long ones.
Let $\tilde{w}_k$ be the number of ways to place the $k^{th}$ $nW$-mer on the surface, given that
$(k-1)$ $nW$-mers have already been placed.  
To begin with, there are $anW-(k-1)nW$ available sites for the first monomer.  Recall that  $q$ is the number of
neighbours of each site in the surface lattice.
In the absence of other polymers, there would be $q$ choices for the second monomer in the chain,
and $q-1$ choices for each monomer after that (here we are using the non-reversed walk model of a polymer
instead of the fully self-avoiding model).  But the number of choices should on average be reduced 
by the fraction of the surface that has already been covered.  
Thus, when we are trying to place the second monomer of our chain, there are
$anW-(k-1)nW-1$ unoccupied sites, so each site has 
probability $[anW-(k-1)nW-1]/[anW]$ of being available.  
Thus there are $q[anW-(k-1)nW-1]/[anW]$ choices for the second monomer.  
Similarly, after $i$ monomers of the current chain have been placed ($i\geq 2$), the fraction of the
surface that has been covered is $[anW-(k-1)nW-i]/[anW]$, so there are 
$(q-1)[anW-(k-1)nW-i]/[anW]$ choices for the $(i+1)^{th}$ monomer in this chain.  
We conclude that 
\begin{eqnarray}
   \tilde{w}_k  & = &   [anW-(k-1)nW]\,\times \,q\left(\frac{ anW-(k-1)nW-1}{anW}\right)    \,\times
   \nonumber    \\
   & &    \hspace{22mm}
     \prod_{i=2}^{nW-1}(q-1)\left(\frac{ anW-(k-1)nW-i}{anW}\right)  
      \nonumber \\
 & = &    \frac{[anW-(k-1)nW]!}{[anW-knW]!} \,q\,  \frac{(q-1)^{nW-2}}{(anW)^{nW-1}}  
     \nonumber    \\
     & = &    \frac{[anW-(k-1)nW]!}{[anW-knW]!}   anW \frac{q}{(q-1)^2}\left(  \frac{q-1}{anW}\right)^{nW}.
     \label{eq.wk2}
\end{eqnarray}
Similarly, let $\tilde{u}_{\ell}$ be the number of ways to place the $\ell^{th}$ $W$-mer on the surface, given that
${\ell}-1$ $W$-mers (or a total of $(\ell-1)W$ monomers) have already been placed.   The same 
argument as for $\tilde{w}_k$ gives
\begin{eqnarray}
   \tilde{u}_{\ell}  
 & = &    \frac{[anW-(\ell-1)W]!}{[anW-\ell W]!} \,q\,  \frac{(q-1)^{W-2}}{(anW)^{W-1}}  
     \nonumber    \\
     & = &    \frac{[anW-(\ell-1)W]!}{[anW-\ell W]!}   anW \frac{q}{(q-1)^2}\left(  \frac{q-1}{anW}\right)^{W}.
     \label{eq.uell2}
\end{eqnarray}

To evaluate $|\mathcal{E}_j|$, we assume that the unadsorbed solution
is dilute enough that we don't need to worry about mutual exclusion between the free floating
polymers.   
For $\mathcal{E}_j$, we have $j$ adsorbed $nW$-mers, $na-nj$ adsorbed $W$-mers, 
$n_L-j$ desorbed $nW$-mers, and $n_S-n(a-j)$ desorbed $W$-mers.   Thus, recalling 
Equations (\ref{eq.sawscale}) and  (\ref{eq.Npoly}), we have
\begin{equation}
    |\mathcal{E}_j|  
      \; = \; \frac{ \Psi_L^{n_L-j} }{(n_L-j)!} \,
          \frac{ \Psi_S^{n_S-n(a-j)} }{(n_S-n(a-j))!} \,
           \frac{\tilde{w}_1\tilde{w}_2\cdots \tilde{w}_j}{j!}
                     \,   \frac{\tilde{u}_{nj+1}\tilde{u}_{nj+2}\cdots \tilde{u}_{an}}{(n(a-j))!}
%                                  \nonumber   \\
%            & & \hspace{7mm}  \times \, \frac{\tilde{w}_1\tilde{w}_2\cdots \tilde{w}_j}{j!}
%                     \,   \frac{\tilde{u}_{nj+1}\tilde{u}_{nj+2}\cdots \tilde{u}_{an}}{(n(a-j))!}  % \,e^{-nW\epsilon/kT}
        \label{eq.Yj}
\end{equation}
where 
\begin{equation}
    \label{eq.psidef}   
   \Psi_L\;=\;V_{latt} A_3 (nW)^{\gamma_3-1} \mu_{3}^{nW}   
    \hspace{5mm}\hbox{and} \hspace{5mm}
     \Psi_S \;=\; V_{latt}A_3W^{\gamma_3-1}\mu_3^W \,.
\end{equation}




To determine which value of $j$ maximizes $|\mathcal{E}_j|$, we look at ratios of consecutive terms:
\begin{eqnarray}
    \frac{|\mathcal{E}_{j+1}|}{|\mathcal{E}_j|}   & = & 
 \frac{
     \left(\frac{\Psi_L^{n_{L}-j-1}}{(n_L-j-1)!}\right)\cdot\left(\frac{\Psi_S^{n_S-n(a-j-1)}}{
    (n_S-n(a-j-1))!}\right) 
      }{
 \left(\frac{\Psi_L^{n_{L}-j}}{(n_L-j)!}\right)\cdot\left(\frac{\Psi_S^{n_S-n(a-j)}}{
    (n_S-n(a-j))!}\right)  }  \,   \frac{\tilde{w}_{j+1} }{  \tilde{u}_{nj+1}\cdots \tilde{u}_{n(j+1)} } 
    \nonumber \\
    & & \hspace{42mm}  
     \times   \, \frac{ j! \, (na-nj)!}{(j+1)!\, (na-nj-n)!}  
    \nonumber    \\
    & = & \frac{\Psi_S^n}{\Psi_L}
   % \frac{ \left(  V_{latt} A_3 W^{\gamma_3-1}\right)^{n-1}}{n^{\gamma_3-1}}
    \,   \frac{\tilde{w}_{j+1} }{  \tilde{u}_{nj+1}\cdots \tilde{u}_{n(j+1)} }   \,\left(  \frac{n_L-j}{j+1}\right)
    \nonumber \\
    & &   \hspace{8mm}  \times \;  \frac{(na-nj)\cdots (na-nj-n+1)}{(n_S-na+nj+1)\cdots(n_S-na+nj+n)} \,.
        \label{eq.Yratio1}
\end{eqnarray}
By Equation (\ref{eq.psidef}), we have
\begin{equation}
   \label{eq.psiratio} 
      \frac{\Psi_S^n}{\Psi_L}   \;=\;   \frac{ (V_{latt}\, A_3 \,W^{\gamma_3-1})^{n-1}}{n^{\gamma_3-1}}\,.
\end{equation}
From Equations (\ref{eq.wk2}) and (\ref{eq.uell2}), we find
\begin{eqnarray}
     \frac{\tilde{w}_{j+1} }{  \tilde{u}_{nj+1}\cdots \tilde{u}_{n(j+1)} }  & = & 
        \left( \frac{(q-1)^2}{qanW}\right)^{n-1}   \times \,
    \frac{   \frac{(anW-jnW)!}{(anW-(j+1)nW)!}  }{
        \prod_{\ell=0}^{n-1} \frac{ (anW-(nj+\ell)W)!}{(anW-(nj+\ell+1)W)!}   }
     \nonumber  \\
     & = & \left( \frac{(q-1)^2}{qanW}\right)^{n-1}  \,\times\, 1.
     \label{eq.Yratioa}     
\end{eqnarray}
Next we use the approximation   $(a+1)(a+2)\cdots (a+m-1) \,\approx \,(a+\frac{m}{2})^m$
(essentially, replacing the geometric mean by the arithmetic mean) to obtain
\begin{equation}
       \frac{(na-nj)\cdots (na-nj-n+1)}{(n_S-na+nj+1)\cdots(n_S-na+nj+n)} 
       \; \approx \;    \frac{ [na-(j+\frac{1}{2})n]^n }{ [n_S-na+nj+\frac{n}{2}]^n} \,.\hspace{3mm}
  %     \nonumber  \\
  %     & \approx & \left(  \frac{n(a-j-\frac{1}{2})}{n_S-(a-j-\frac{1}{2})}\right)^n  .
      \label{eq.Yratiob}
\end{equation} 
Now, putting Equations (\ref{eq.psiratio}--\ref{eq.Yratiob}) back into (\ref{eq.Yratio1}), we obtain
\begin{equation}
    \label{eq.Yratio2}
       \frac{|\mathcal{E}_{j+1}|}{|\mathcal{E}_j|} \; \approx \; 
       \left(  \frac{ V_{latt}A_3W^{\gamma_3-1}(q-1)^2}{aWq} \,
          \frac{(a-j-\frac{1}{2})}{n_S-n(a-j-\frac{1}{2})} \right)^{n+o(n)}   \,.
\end{equation}
The quantity on the right of Equation (\ref{eq.Yratio2}) decreases in $j$.  Let $j^*$ be the 
smallest value of $j$ for which the right-hand side of Equation (\ref{eq.Yratio2}) is less than 1.
Then we see that $|\mathcal{E}_{j+1}|/|\mathcal{E}_j|<1$ whenever $j\geq j^*$, and 
$|\mathcal{E}_{j+1}|/|\mathcal{E}_j|\geq 1$ whenever $j< j^*$.
That is, $|\mathcal{E}_j|$ is decreasing in $j$ when $j\geq j^*$, and 
increasing when $j< j^*$.
It follows that $|\mathcal{E}_j|$ is maximized at $j=j^*$, and that we can find $j^*$ by seeing 
where the right-hand side of Equation (\ref{eq.Yratio2}) is close to 1.   Thus $j^*$ satisfies
\[     1  \;\approx \; 
     K  \,%\frac{ V_{latt}A_3W^{\gamma_3-1}(q-1)^2}{aWq} \,
          \frac{a-j^*-\frac{1}{2}}{n_S-n(a-j^*-\frac{1}{2})}    \,,  
             \hspace{5mm}\hbox{where}\hspace{5mm}
             K \;=\;   \frac{ V_{latt} }{aW}\, \frac{   A_3W^{\gamma_3-1}(q-1)^2}{q} \,,
\]
which can be rewritten as 
\[       a-j^*-\frac{1}{2} \; \approx\;    \frac{n_S}{K+n}   \,.
\]
We observe that $K\gg n$; indeed,  $V_{latt}/aWn \,=\,V_{latt}/A_{latt}\gg 1$.  
Thus we can replace $K+n$ by $K$ in the above, and omit the term $\frac{1}{2}$, resulting in the   
approximation 
\begin{equation}
    \label{eq.jstar}
     1-\frac{j^*}{a}     \; \approx   \; 
        \left(  \frac{n_S\,W}{V_{latt} }\right) \,\left(   \frac{q}{A_3W^{\gamma_3-1}(q-1)^2}\right)  \,.
\end{equation}
The above equation can be interpreted as follows.  The left hand side, $1-j^*/a$, is the fraction of the
sphere's surface that is covered by short monomers.  The ratio $n_SW/V_{latt}$ gives
the concentration of monomers corresponding to those appearing on the short chains only.
The quantity in the second set of 
parentheses in Equation (\ref{eq.jstar}) cannot be large, since $q<(q-1)^2$ and $\gamma_3>1$.
If the solution is fairly dilute, then we see that the left side of (\ref{eq.jstar}) is small,
and hence only a small fraction of the surface is covered by short polymers.


[\textit{Remark:}  Observe that $n_L$ does not appear in Equation (\ref{eq.jstar}).  
In the expression for $|\mathcal{E}_{j+1}|/|\mathcal{E}_j|$,   only $n_S$ appears inside the 
expression that is raised to the power $n$, while $n_L$ appears outside this expression and is
raised to the power 1.]  



\section{One-dimensional model}

We now present a more tractable ``toy'' version of the lattice model.  We shall assume that 
the surface is one-dimensional, which eliminates the need for the Flory-Huggins-type approximations.

As before, we have $n_S$ short $W$-mers and $n_L$ long $nW$-mers, some of which are adsorbed 
onto a surface with $anW$ lattice sites, and the rest float as free self-avoiding walks 
in a three-dimensional cubic lattice with  $V_{latt}$ sites.  
The difference now is that we shall replace the spherical surface with a line segment
$\mathcal{S}$ containing $anW$ sites.  When a short (respectively, long) polymer is adsorbed, it covers a 
sub-segment of $\mathcal{S}$ containing $W$ (respectively, $nW$) sites.
If exactly $j$ long polymers are adsorbed onto $\mathcal{S}$, then $n(a-j)$ short polymers are
also adsorbed.  The number of ways that this can be done depends only on the order in which 
longs and shorts appear when traversing $\mathcal{S}$ from one end to the other.  Since the total 
number of adsorbed polymers in this case is $n(a-j)+j$, the number of ways is 
\[      \binom{n(a-j)+j}{j}   \,.  \]
As before, let $\mathcal{E}_j$ be the number of ways to place $j$ $nW$-mers and $a(n-j)$ $W$-mers on 
$\mathcal{S}$, as well as $n_L-j$ $nW$-mers and $n_S-a(n-j)$ $W$-mers as self-avoiding walks 
in the three dimensional lattice.  Then we have 
\begin{equation}
   \label{eq.1dE}
    |\mathcal{E}_j| \;=\;  \left(\frac{\Psi_L^{n_{L}-j}}{(n_L-j)!}\right)\cdot\left(\frac{\Psi_S^{n_S-n(a-j)}}{
    (n_S-n(a-j))!}\right) \cdot\binom{an-jn+j}{j}   \,,
\end{equation}
where 
$\Psi_L$ and $\Psi_S$ are as defined in Equation (\ref{eq.psidef}).
%\[       \Psi_L\;=\;V_{latt} A_3 (nW)^{\gamma_3-1} \mu_{3}^{nW}   
%    \hspace{5mm}\hbox{and} \hspace{5mm}
%     \Psi_S \;=\; V_{latt}A_3W^{\gamma_3-1}\mu_3^W \,.    \]

We shall show that for the one-dimensional model, $|\mathcal{E}_j|$ increases rapidly as $j$ increases,
and hence the likeliest situation is complete coverage by long polymers.  To do this, we shall show that
$|\mathcal{E}_{j+1}|/|\mathcal{E}_j| \,\gg\,1$ for every $j=0,\ldots,a-1$.

From Equation (\ref{eq.1dE}), we have
\begin{align}
   \nonumber
    \frac{|\mathcal{E}_{j+1}|}{|\mathcal{E}_j|} \; & =\;   
    \frac{
     \left(\frac{\Psi_L^{n_{L}-j-1}}{(n_L-j-1)!}\right)\cdot\left(\frac{\Psi_S^{n_S-n(a-j-1)}}{
    (n_S-n(a-j-1))!}\right) \cdot\left( \frac{ (an-jn-n+j+1)!}{(j+1)!(an-jn-n)!} \right)
      }{
 \left(\frac{\Psi_L^{n_{L}-j}}{(n_L-j)!}\right)\cdot\left(\frac{\Psi_S^{n_S-n(a-j)}}{
    (n_S-n(a-j))!}\right) \cdot   \left( \frac{ (an-jn+j)!}{j!(an-jn)!} \right)    }
    \\
    \nonumber
    & = \;   \left( \frac{\Psi_S^n \,(n_L-j) }{\Psi_L \, \prod_{k=1}^n[n_S-na+nj+k]  }\right)  \,\times 
     \\
     \nonumber
    & \hspace{31mm}  \left(   \frac{  \prod_{k=0}^{n-1}[an-jn-k]    }{(j+1)  \prod_{k=0}^{n-2} [an-jn+j-k]   }\right)
    \\
    \label{eq.1dEE}
      & \approx \;   \frac{ \Psi_S^n \,(n_L-j) [an-jn-\frac{n}{2}]^n
         }{   \Psi_L (j+1)  [n_S-na+nj+\frac{n}{2}]^n  [an-jn+j-\frac{n}{2}]^{n-1} }   \,.
\end{align}
Using Equation (\ref{eq.psiratio}) in 
 Equation (\ref{eq.1dEE}), we obtain 
\begin{equation}
   \label{eq.1dEEb}  
    \frac{|\mathcal{E}_{j+1}|}{|\mathcal{E}_j|} \; \approx  \;   
      \left(   \frac{ V_{latt}\, A_3 \,W^{\gamma_3-1} }{n_S-na+nj+\frac{n}{2}} \cdot 
         \frac{ an-nj-\frac{n}{2} }{an-jn-\frac{n}{2}+j}   \right)^{n+o(n)}.
\end{equation}
The first quotient inside the parentheses in Equation (\ref{eq.1dEEb}) is large:  indeed,
$V_{latt}/(n_S-na+nj)$ equals $W$ divided by the fraction of the solvent lattice that is occupied by 
short polymers.  
And the second quotient inside the parentheses is close to 1 for large $n$.   Therefore 
Equation (\ref{eq.1dEEb}) tells us that $|\mathcal{E}_{j+1}|/|\mathcal{E}_j| \,\gg\,1$
for every $j$.  It follows that configurations are highly likely to be in $\mathcal{E}_a$, i.e.\
to have $\mathcal{S}$ covered entirely by long polymers.


\begin{thebibliography}{99}

\bibitem{CL}  M. Chen and K.Y. Lin (2002), ``Universal amplitude ratios for three-dimensional
self-avoiding walks.''  \textit{J. Phys.\ A:  Math.\ Gen.} \textbf{35}, 1501--1508.

%\bibitem{DBL}  K. De'Bell and T. Lookman (1983), ``Surface phase transitions in polymer systems.''
% \textit{Reviews of Modern Physics}, vol.\ 65, no.\ 1, 87--113.

\bibitem{Flo}  P.J. Flory (1953), \textit{Principles of Polymer Chemistry}.  Cornell University Press, Ithaca.

\bibitem{MS}  N. Madras and G. Slade (1993), \textit{The Self-Avoiding Walk}.  Birkh\"{a}user, Boston.


\end{thebibliography}


\end{document}










%%%%%%%%%%  CUT:  




\section{First approach:  Counting self-avoiding walks}

To convey what I think is the main idea,
consider the case that we have one long linear polymer in solution as well as several 
short polymers whose combined size equals that of the long polymer.
That is, 
suppose we have $n+1$ linear polymers in solution, of which $n$ have $W$ monomers and 
one has $nW$ monomers.   The example in Section 6 of the draft has $W=100$ and $n=3$;
the experiment described in the draft corresponds to $n= 450K/2K = 225$.
We wish to compare the following two scenarios:
\begin{verse}
(L):  The \underline{L}ong $(nW)$-mer is adsorbed and the $n$ (short) $W$-mers are free;
\\
(S):   The $n$ \underline{S}hort $W$-mers are adsorbed and the (long) $nW$-mer is free.
\end{verse}
%To make the notation a bit lighter, I'll assume $k=3$, in accordance with the example in Section 6.

We shall model the polymers as self-avoiding walks (SAWs)  of lengths 
$W$ and $nW$ in the simple cubic lattice.
(The choice of lattice will not matter much.)   A fundamental property of SAWs is the following.  
Let $c_{\ell}$ be
the number of ${\ell}$-step SAWs that start from a specified lattice site (``the origin'').   Then the 
asymptotic behaviour of $c_{\ell}$ is
\begin{equation}
    \label{eq.sawscale}
       c_{\ell}  \;\sim  \;  A \, {\ell}^{\gamma-1}  \mu^{\ell}    \hspace{5mm}\hbox{as $\ell\rightarrow\infty$}.
\end{equation}
(This has not been proven rigorously, but even mathematicians do not doubt its truth.) 
Here $\mu$ and $A$ depend on the details of the lattice, but $\gamma$ is a universal critical exponent
that only depends on the dimension of the lattice.  Free polymers in solution are modelled
by SAWs on a three-dimensional lattice.  In contrast, a polymer that is completely adsorbed onto a
flat surface is modelled by a SAW on a \textit{two}-dimensional lattice.  
So let us write $c^{(3)}_{\ell}$, $A_3$, $\gamma_3$, and $\mu_3$ to denote these quantities on the 
(three-dimensional) simple cubic lattice, and  
$c^{(2)}_{\ell}$, $A_2$, $\gamma_2$, and $\mu_2$ for the 
corresponding quantities on the (two-dimensional) square lattice.  
As mentioned above, the values of $\gamma_3$ and $\gamma_2$ depend only on the dimension
of the corresponding lattice.  Their values are known to be
\begin{equation}
   \label{eq.gammas}
        \gamma_3 \;=\;  1.162\ldots    \hspace{5mm}\hbox{and}\hspace{5mm}
    \gamma_2  \;=\;  \frac{43}{32}   \;=\;   1.34375  \,.
\end{equation}

\smallskip
To make the argument simpler at this point, we shall assume that the adsorbed polymer lies 
entirely in the surface.  (I shall explain later how we can circumvent this assumption.)
I shall generally follow the notation currently in Section 6 of the draft.  We shall neglect each polymer's 
interactions with the other $n$ polymers.

Under the assumption that exactly one of the scenarios (L) or (S) occurs, 
the partition function can be written as 
\[    Z   \;=\;  Z_L \,+\,Z_S  \,    \]
where $Z_L$ and $Z_S$ correspond to the scenarios (L) and (S).  In case (S),
the desorbed $nW$-mer contributes $Z_{\textrm{vol}}c^{(3)}_{nW}$ to the partition function 
(as in equation (1) 
in the draft), while each adsorbed $W$-mer contributes $Z_{\textrm{surf}}c_{W} \exp(-W\epsilon/kT)$  (as in 
equation (3) in the draft; my assumption of complete adsorption into the surface means that the 
sum in equation (3) of the draft omits every term except $n=L$).  I'll say more about $Z_{\textrm{vol}}$ 
and $Z_{\textrm{surf}}$ below.
Thus we have
\begin{equation}
   \label{eq.ZS}
      Z_S  \;=\;  \left(Z_{\textrm{vol}}\,c^{(3)}_{nW}\right)\left(  Z_{\textrm{surf}}\,c^{(2)}_{W} \,
         e^{-W\epsilon/kT} \right)^n  \,.
        %\exp(-N\epsilon/kT) \right)^3  \,.
\end{equation}
An analogous argument for case (L) gives
\begin{equation}
   \label{eq.ZL}
      Z_L  \;=\;  \left(Z_{\textrm{vol}}\,c^{(3)}_{W}\right)^n\left(  Z_{\textrm{surf}}\,c^{(2)}_{nW} \,
         e^{-nW\epsilon/kT} \right)  \,.
        %\exp(-N\epsilon/kT) \right)^3  \,.
\end{equation}
Now we need to compare $Z_S$ with $Z_L$.  First notice that both have the same total energy term,
$nW\epsilon$, so any difference will be entirely due to entropy.  We have
\begin{eqnarray}
\frac{Z_L}{Z_S}  & = &    \frac{  \left(Z_{\textrm{vol}}\,c^{(3)}_{W}\right)^n\left(  Z_{\textrm{surf}}\,c^{(2)}_{nW} \,
         e^{-nW\epsilon/kT} \right) }{   \left(Z_{\textrm{vol}}\,c^{(3)}_{nW}\right)\left(  Z_{\textrm{surf}}\,c^{(2)}_{W} \,
         e^{-W\epsilon/kT} \right)^n }   
         \nonumber \\
         & \sim & \left(\frac{ Z_{\textrm{vol}} }{Z_{\textrm{surf}} }\right)^{n-1}  
         \frac{ \left(A_3 W^{\gamma_3-1}\mu_3^W\right)^n \left(A_2 (nW)^{\gamma_2-1}\mu_2^{nW} \right)}{
         \left(A_3(nW)^{\gamma_3-1}\mu_3^{nW}\right) \left(A_2 W^{\gamma_2-1}\mu_2^{W}\right)^n }
         \nonumber \\
         & = & 
         \left(\frac{ Z_{\textrm{vol}}\,A_3 \,W^{\gamma_3-1}}
         {Z_{\textrm{surf}}\, A_2  \,W^{\gamma_2-1}}\right)^{n-1} \, n^{\gamma_2-\gamma_3}
         \nonumber \\
         & = &   \left(\frac{ Z_{\textrm{vol}}\,A_3 }{Z_{\textrm{surf}}\, A_2}\, W^{-0.18}\right)^{n-1} \,
            n^{0.18}     
      \label{eq.ZZ}
\end{eqnarray}
where in the last step we have plugged in the approximate values $\gamma_2=1.34$ and $\gamma_3=1.16$
from Equation (\ref{eq.gammas}).

There are different effects at play in Equation (\ref{eq.ZZ}).  Observe that the ratio $Z_L/Z_S$ decreases
as $W$ gets large, but it increases as $n$ gets large.  $A_3$ and $A_2$ are constants.  The major
difference, I think, is the ratio $Z_{\textrm{vol}}/Z_{\textrm{surf}}$, which will be large assuming that 
the surface area where adsorption occurs is small in comparison to the total volume. 

As mentioned in Section 6 of the draft, $Z_{\textrm{vol}}$ is the translational partition function 
within the solution volume, and $Z_{\textrm{surf}}$ is the translational partition function 
on the adsorbing surface.
My interpretation (which may not be quite right) is the following: 
If we think of the solution region as being covered by a fine lattice, then $Z_{\textrm{vol}}/Z_{\textrm{surf}}$
is something like the ratio of the number of lattice points that lie in the solution volume to the number
that lie close to the adsorbing surface.

The above argument was based on the assumption that the adsorbed polymer lies completely in the surface.
This is not generally true unless $|\epsilon/kT|$ is very large.  Instead, Equation (3) from the draft
says that the contribution of an adsorbed SAW of length $\ell$ is 
\[     Z_{\textrm{surf}}\, z_{\ell}(\epsilon/kT),    \hspace{5mm}  \hbox{where}\hspace{5mm}
    z_{\ell}(\omega)  \;=\;   \sum_{j=1}^{\ell}N_{j,\ell}\,e^{-j\omega}
\]
(here, $N_{j,\ell}$ is the number of different ${\ell}$-step SAWs such that exactly $j$ sites contact the 
surface).  As noted in Section III(A) of \cite{DBL}, when the system is above the adsorption threshold
we have 
%(writing $\omega$ for $-\epsilon/kT$)
\[      z_{\ell}(\omega)   \;\sim\;   A_2(\omega) \, \ell^{\gamma_2-1} h(\omega)^{\ell}   \]
for some function $h$.  Notably, the exponent $\gamma_2$ is the same as in Equation (\ref{eq.gammas}).
So the calculations proceed as before, with $\mu_2$ replaced by $h(\omega)$, and we see that 
$h(\omega)$  cancels out in the end, exactly as $\mu_2$ did.

\medskip
In conclusion:  I claim that $Z_L$ is much larger than $Z_S$, due mainly to the fact that 
$Z_{\textrm{vol}}$ is significantly larger than $Z_{\textrm{surf}}$.
Therefore Scenario (L) is much more probably than Scenario (S).   If this argument is
correct, then my feeling is that an extension should also work for systems where we have many 
long and short polymers.

%%%%%%%%%%%%%%%%%


First, there are $anW-(k-1)nW$ available sites for the first monomer.  Recall that  $q$ is the number of
neighbours of each site in the surface lattice.
In the absence of other polymers, there would be $q$ choices for the second monomer in the chain,
and $q-1$ choices for each monomer after that (here we are using the non-reversed walk model
instead of the fully self-avoiding model).
However, because of the previously placed polymers, we expect that each site has a 
probability $[anW-(k-1)nW]/[anW]$ of being available.  
Thus there are $q[anW-(k-1)nW]/[anW]$ choices for the second monomer, and 
$(q-1)[anW-(k-1)nW]/[anW]$ choices for each subsequent monomer in this chain.  
We conclude that 
\begin{eqnarray}
   w_k  & = &    [anW-(k-1)nW] \,q(q-1)^{nW-2}\left(  \frac{anW-(k-1)nW}{anW}\right)^{nW-1}
     \nonumber    \\
     & = &    anW \frac{q}{(q-1)^2}\left(  \frac{(q-1)(a-(k-1))}{a}\right)^{nW}.
     \label{eq.wk1}
\end{eqnarray}

Similarly, let $u_{\ell}$ be the number of ways to place the $\ell^{th}$ $W$-mer on the surface, given that
${\ell}-1$ $W$-mers (or a total of $(\ell-1)W$ monomers) have already been placed.   The same 
argument as for $w_k$ gives
\begin{eqnarray}
   u_{\ell}  & = &    [anW-(\ell-1)W] \,q(q-1)^{W-2}\left(  \frac{anW-(\ell-1)W}{anW}\right)^{W-1}
     \nonumber    \\
     & = &    anW \frac{q}{(q-1)^2}\left(  \frac{(q-1)(an-(\ell-1))}{an}\right)^{W}.
     \label{eq.uell1}
\end{eqnarray}





